Canonical rings of $\bq$-divisors on Minimal Rational Surfaces

### 2015-08-17 surfaces-write-up/surface-write-up.tex
respond to peters comments like misuse of m for indexing lambda

\maketitle
\todo{Aaron: Remove labels which are not referenced
\\
Peter: I did this for a few of them, but I'm wondering what the purpose of this is.  It appears the same in the PDF either way right?}
\todo{Aaron: Make sure indexing at 0 is correct in projective space
section, particularly toward the end}
%%%%%%%%%%%%%%%%%%%%%%%%%%%% Introduction %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

In Section \ref{sec:proj} we prove the following two results bounding
the degree of generators and relations of
stacky canonical rings on $\bp^m$:
\begin{restatable}{thm}{restateeffective}
\label{thm:proj-effective-intro}
Let $D = \sum_{i = 0}^{n} \alpha_i D_i \in \bz \otimes_\bz \di \bp^m$, with $\alpha_i =
\frac{c_i}{k_i} \in \bq_{> 0}$ and let each $D_i$ be a hypersurface of
$\bp^m$.

@@

\noindent
is generated in degrees at most $\max_{0 \leq i \leq n}{k_i}$ with
relations generated in degrees at most $2 \max_{0 \leq i \leq n}{k_i}$
\end{restatable}

@@

\ssec{Notations and Conventions}
We will use $m \in \bz$ to index the dimension of a given projective
space $\bp^m$ or the type of the hirzebruch surface $\hirz_m$.
Let $D$ be a (Weil) $\bq$-divisor of the form
\begin{align*}
	D = \sum_{i=1}^{n}\alpha_i D_i \in \di X \otimes \bq.
\end{align*}\todo{Peter: Do we want this indexing from 0 or 1, since we use 0 for $\bp^m$ and $1$ for $\mathbb{F}_m$.}

@@

\begin{proof}
By definition of $H^0(\bp^m,dD)$, functions satisfying conditions 
~\eqref{align:pm-span} lie in $H^0(\mathbb{P}^m,dD)$, and all functions of $H^0(\mathbb{P}^m, dD)$ are of that form.
\todo{Aaron: Peter, what do you mean when you say ``all functions of that form lie in...''? How is
this different from saying that functions satisfying that equation lie in $H^0$?
\\
Peter:  Oops, meant to write the converse statement, but accidently just restated the original statement-it's fixed now.}
To complete the proof, it
suffices to check functions satisfying conditions ~\eqref{align:pm-basis} form
a basis of $H^0(\bp^m,dD)$. Note that there are $\binom{m+ \deg \lfloor dD \rfloor }{m}$
functions satisfying condition ~\eqref{align:pm-basis}. However we know
$h^0(\bp^m,dD) = \binom{m+ \deg \lfloor dD \rfloor }{m},$ so it suffices to show
that those functions satisfying condition ~\eqref{align:pm-basis} are
independent. This follows from the assumption that $f_0,\ldots f_m$ form a
basis of linear forms, and so all monomials in $f_0,\ldots, f_m$
of degree $\deg \lfloor d \deg D \rfloor $
form a basis of degree $\deg \lfloor d  D \rfloor$ rational functions.
\end{proof}

@@

\sssec*{Idea of Proof:}
The proof follows in four steps. First, we demonstrate that
the $F_{i}^{\vec{v}}$'s form a set of minimal generators of $R_D$
over $R_{D'}$ using their construction in terms of best lower
approximations to $\alpha$. Then, since the $F_{i}^{\vec{v}}$ generate
all of $R_D$ over $R_{D'}$, we obtain the
relations $G_{i, j}^{\vec{v}, \vec{w}}$ \todo{Peter: hmm... this doesn't say much about how we get the $G_{i,j}$'s.}.
We derive the relations
$L_{i, j}^{\vec{v}, \vec{w}}$ by considering when products of
generators in neighboring degrees are equal. Finally, we demonstrate
that $G_{i, j}^{\vec{v}, \vec{w}}$'s and $L_{i, j}^{\vec{v}, \vec{w}}$'s
generate all of the relations by using them to reduce arbitrary
elements of
$R_D$ to a canonical form.

@@

\begin{lem}
\label{lem:bound-ker-chi}
Retaining the notation of Lemma ~\ref{lem:composite-map}, if $\Sigma$ has
extremal rays $e_0,\ldots, e_n$ in degrees $d_0, \ldots, d_n$ then $\ker \chi$
is generated in degrees at most $2(\sum_{i=0}^{n}d_i-1)$.
\end{lem}
\begin{proof}
Since $e_0, \ldots, e_n$ are extremal rays, by Proposition
~\ref{prop:cone-generation}, every element $\sigma \in \Sigma$ can be written
in a canonical form $\lambda + \sum_{i=0}^{n} r_i e_i$. Now, let $\lambda_0 \colon =
0,\lambda_1, \ldots, \lambda_m$ be all elements of $\Sigma$ which can be
written in the form $\lambda_i = \sum_{i=0}^{n}r_i e_i$ with $0 \leq r_i < 1.$ \todo{Peter: This sentence needs rephrasing.  1: misuse of $i$ as index.  2: misuse of $m$ as index ($\bp^m$).}
Then, for any $1 \leq i \leq j \leq m,$ we can write $\lambda_i + \lambda_j$ in
the above canonical form, yielding a relation in degree at most $\deg \lambda_i
+ \deg \lambda_j \leq 2 \cdot \left( \sum_{i=0}^{n}d_i -1 \right).$
Furthermore, these relations generate all relations, as one can apply a
sequence of these relations to put any $\sigma \in \Sigma$ into canonical form
$\sigma = \lambda + \sum_{i=0}^{n}w_i e_i$ as in Equation
~\eqref{eqn:sigma-canonical-form}.
\end{proof}
\begin{lem}
\label{lem:proj-relations-psi}
Let $D = \sum_{i=0}^{n} \alpha_i D_i \in \bq \otimes \di \bp^m,$ where
$\deg D_i = \bida_i$, $\alpha_i = \frac{c_i}{k_i}\in \bq$, and
$\ell_i = \lcm_{j \neq i} (k_j)$. Define $\Sigma$ as in Equation
~\eqref{eqn:Sigma-defn} and $\psi$ as ~\eqref{eqn:factor-through-semigroup-ring}. Then, $\ker \psi$ is generated in degrees at most
\begin{align}
\label{eqn:proj-relation-degree}
	\frac{\max_{0\le i \le n}(\bida_i)}{\deg(D)} +  \sum_{i=0}^n \ell_i a_i.
\end{align}

>>>>

\maketitle
\todo{Aaron: Remove labels which are not referenced
\\
Peter: I did this for a few of them, but I'm wondering what the
purpose of this is.  It appears the same in the PDF either way
right?

Aaron: Well, the point is to actually remove the references
numbers from equations which aren't referenced, since this
removes clutter. That is, put a star after the environment
name (change begin{equation} to begin{equation*}).
I think it makes sense not to label things
if you don't use those labels, and 
David recommended doing this in his comments.
}
\todo{Aaron: Make sure indexing at 0 is correct in projective space
section, particularly toward the end}
%%%%%%%%%%%%%%%%%%%%%%%%%%%% Introduction %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

In Section \ref{sec:proj} we prove the following two results bounding
the degree of generators and relations of
stacky canonical rings on $\bp^m$:
\begin{restatable}{thm}{restateeffective}
\label{thm:proj-effective-intro}
Let $D = \sum_{i = 0}^{n} \alpha_i D_i \in \bz \otimes_\bz \di \bp^m$, with $\alpha_i =
\frac{c_i}{k_i} \in \bq_{> 0}$ and let each $D_i \in \di \bp^m$ be an integral divisor.

@@

\noindent
is generated in degrees at most $\max_{0 \leq i \leq n}{k_i}$ with
relations generated in degrees at most $2 \max_{0 \leq i \leq n}{k_i}$.
\end{restatable}

@@

\ssec{Notations and Conventions}
We will use $m \in \bz$ to index the dimension of a given projective
space $\bp^m$ or the type of the hirzebruch surface $\hirz_m$.
Let $D$ be a (Weil) $\bq$-divisor of the form
\begin{align*}
	D = \sum_{i=1}^{n}\alpha_i D_i \in \di X \otimes \bq.
\end{align*}\todo{Peter: Do we want this indexing from 0 or 1,
	since we use 0 for $\bp^m$ and $1$ for $\mathbb{F}_m$.

Aaron: Read the next two sentences. I think 1 is fine.}

@@

\begin{proof}
By definition of $H^0(\bp^m,dD)$, functions satisfying conditions 
~\eqref{align:pm-span} lie in $H^0(\mathbb{P}^m,dD)$.
\todo{Aaron: Peter, what do you mean when you say ``all functions of that form lie in...''? How is
this different from saying that functions satisfying that equation lie in $H^0$?
\\
Peter:  Oops, meant to write the converse statement, but accidently
just restated the original statement-it's fixed now.

Aaron: Well, it should really be all such functions are linear
combinations of things in that form. However,
that's what I proved in the next couple setences. I.e.,
the remainder of the proof is showing that this set is indeed
a basis. I removed your added clause.
}
To complete the proof, it
suffices to check functions satisfying conditions ~\eqref{align:pm-basis} form
a basis of $H^0(\bp^m,dD)$. Note that there are $\binom{m+ \deg \lfloor dD \rfloor }{m}$
functions satisfying condition ~\eqref{align:pm-basis}. However we know
$h^0(\bp^m,dD) = \binom{m+ \deg \lfloor dD \rfloor }{m},$ so it suffices to show
that those functions satisfying condition ~\eqref{align:pm-basis} are
independent. This follows from the assumption that $f_0,\ldots f_m$ form a
basis of linear forms, and so all monomials in $f_0,\ldots, f_m$
of degree $\deg \lfloor d \deg D \rfloor $
form a basis of degree $\deg \lfloor d  D \rfloor$ rational functions.
\end{proof}

@@

\sssec*{Idea of Proof:}
The proof follows in four steps. First, we demonstrate that
the $F_{i}^{\vec{v}}$'s form a set of minimal generators of $R_D$
over $R_{D'}$ using their construction in terms of best lower
approximations to $\alpha$. Then, since the $F_{i}^{\vec{v}}$ generate
all of $R_D$ over $R_{D'}$, and the leading terms of
$G_{i,j}^{\vec v, \vec w}$ lies in $R_{D'}$, we obtain the
relations $G_{i, j}^{\vec{v}, \vec{w}}$. We derive the relations
$L_{i, j}^{\vec{v}, \vec{w}}$ by considering when products of
generators in neighboring degrees are equal. Finally, we demonstrate
that $G_{i, j}^{\vec{v}, \vec{w}}$'s and $L_{i, j}^{\vec{v}, \vec{w}}$'s
generate all of the relations by using them to reduce arbitrary
elements of
$R_D$ to a canonical form.

@@

\begin{lem}
\label{lem:bound-ker-chi}
Retaining the notation of Lemma ~\ref{lem:composite-map}, if $\Sigma$ has
extremal rays $e_0,\ldots, e_n$ in degrees $d_0, \ldots, d_n$ then $\ker \chi$
is generated in degrees at most $2(\sum_{i=0}^{n}d_i-1)$.
\end{lem}
\begin{proof}
Since $e_0, \ldots, e_n$ are extremal rays, by Proposition
~\ref{prop:cone-generation}, every element $\sigma \in \Sigma$ can be written
in a canonical form $\lambda + \sum_{i=0}^{n} r_i e_i$. Now, let
$\lambda_0 \colon =
0,\lambda_1, \ldots, \lambda_t$ be all elements of $\Sigma$ which
can be
written in the form $\lambda_i = \sum_{i=0}^{n}r_i e_i$ with $0
\leq r_i < 1.$ Then, for any $1 \leq j \leq k \leq t,$ we can write $\lambda_j + \lambda_k$ in
the above canonical form, yielding a relation in degree at most $\deg \lambda_j
+ \deg \lambda_k\leq 2 \cdot \left( \sum_{i=0}^{n}d_i -1 \right).$
Furthermore, these relations generate all relations, as one can apply a
sequence of these relations to put any $\sigma \in \Sigma$ into canonical form
$\sigma = \lambda + \sum_{i=0}^{n}w_i e_i$ as in Equation
~\eqref{eqn:sigma-canonical-form}.
\end{proof}
\begin{lem}
\label{lem:proj-relations-psi}
Let $D = \sum_{i=0}^{n} \alpha_i D_i \in \bq \otimes \di \bp^m,$ where
$\deg D_i = \bida_i$, $\alpha_i = \frac{c_i}{k_i}\in \bq$, and
$\ell_i = \lcm_{j \neq i} (k_j)$. Define $\Sigma$ as in Equation
~\eqref{eqn:Sigma-defn} and $\psi$ as ~\eqref{eqn:factor-through-semigroup-ring}. Then, $\ker \psi$ is generated in degrees at most
\begin{align}
\label{eqn:proj-relation-degree}
	\frac{\max_{0\le i \le n}(\bida_i)}{\deg(D)} +  \sum_{i=0}^n \ell_i a_i.
\end{align}
